---
title: "Data visualization tools"
author:
  - [TEAM NAME GOES HERE]
output: 
  html_document:
    df_print: paged
    highlight: textmate 
    theme: readable
    code_folding: hide
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
---

# Overview

These exercises will help you practice applying `pivot_wider` and `pivot_longer`, and will also require some use of skills you have previously learned. No new data (other than toy examples) will be introduced.

## Setup

```{r, message = FALSE, warning = FALSE}

# set global rmarkdown options
# echo = TRUE     --> show code blocks in markdown doc
# message = FALSE --> suppress package startup and other messages
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# I am not loading packages this time.
# You can load your own as you determine what tools are needed

```

# Pre-requisites

Before starting these exercises, you should have a good understanding of 

1. The [Tidy your data](https://rstudio.cloud/learn/primers/4.1) Primer.

1. Chapter 12.1 - 12.3 of [R for Data Science](https://r4ds.had.co.nz/tidy-data.html#introduction-6)

# Exercise 1

Convert the given dataset from wide to long form

```{r}

preg <- tibble::tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

```

Your solution should look like this:

```{r}

readr::read_rds('solutions/01_solution.rds')

```

# Exercise 2

Convert the long dataset you created in exercise 1 into a wide dataset with 

- `sex` a column indicating male or female sex.

- `preg_yes` count of pregnancies for the given sex

- `preg_no` count of non-pregnancies for the given sex.

```{r}

# Your code goes here

```

Your solution should look like this:

```{r}

readr::read_rds('solutions/02_solution.rds')

```

# Exercise 3

Convert the following data into long form, with one column for `sbp`, `dbp`, and `hr` (heart rate). 

```{r}
bp <- tibble::tribble(
  ~subjid, ~sbp_1, ~sbp_2, ~dbp_1, ~dbp_2, ~hr_1, ~hr_2,
  'a',      123,    NA,    68,     72,     NA,    70,    
  'b',      140,    139,    NA,     82,     NA,    65
)
```


```{r}

# Your code goes here

```


Your solution should look like this:

```{r}

readr::read_rds('solutions/03_solution.rds')

```

# Exercise 4

Using the data you created in exercise 3, add three new columns that measure the number of valid `sbp`, `dbp`, and `hr` measurements for each `subjid`.

```{r}

# Your code goes here

```

Your solution should look like this:

```{r}

readr::read_rds('solutions/04_solution.rds')

```

# Exercise 5

- Read in the `nhanes_long.rds` file, located in the `data` directory.

- Pivot the data into a wider format and create the table shown below.

```{r}

# Your code goes here

```


```{r}

readr::read_rds('solutions/05_solution.rds')

```

